What is an Ancestral Recombination Graph?

### 2022-02-09 paper.tex
Put figure 1 in the right place

\begin{figure}
\caption{\label{fig-sim-statespace} An illustration of the state space
of Hudson's algorithm (A) and the big ARG (B) at a given instant. The
state maintained by Hudson's algorithm is considerably more complicated,
but is guaranteed to only simulate events corresponding to genetic
ancestors of the sample.
}
\end{figure}

@@

We describe here an alternative formulation in which the
concepts are based on the diploid pedigree, which
is free from the limitations discussed in the previous section.
\cite{mathieson2020ancestry} and others 
\citep{wakeley2012genetics,gusfield2014recombinatorics,speed2015naturereviewsgenetics}
have also suggested this approach, although without explicitly defining
the roles of nodes and the details of how ancestry
flows along edges.

\begin{figure}[ht]
\centering
\begin{tikzpicture}
	% Samples at x = 0, 4, and 6
	\node [circle, fill, scale=0.2,label=above right:{\small A}] at (0,0) {A};
	\draw (-0.4, -0.2) -- (0.4, -0.2) -- (0.4, -0.3) -- (-0.4, -0.3) -- (-0.4, -0.2);
	\draw (-0.1, -0.1) -- (-0.1, -0.4);
	\node [label=below:{\small $v$}] at (-0.1, -0.2) {};
	\node [circle, fill, scale=0.2,label=above right:{\small B}] at (4,0) {B};
	\draw (3.6, -0.2) -- (4.4, -0.2) -- (4.4, -0.3) -- (3.6, -0.3) -- (3.6, -0.2);
	\node [circle, fill, scale=0.2,label=above left:{\small C}] at (6,0) {C};
	\draw (5.6, -0.2) -- (6.4, -0.2) -- (6.4, -0.3) -- (5.6, -0.3) -- (5.6, -0.2);

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (11.1, 1.2); % D
	\draw[color=gray, dashed] (7.5, 1.5) -- (11.1, 1.5); % E and F
	\draw[color=gray, dashed] (7.5, 2.4) -- (11.1, 2.4); % G and H
	\draw[color=gray, dashed] (7.5, 3.1) -- (11.1, 3.1); % I and J
	\draw[color=gray, dashed] (7.5, 4.5) -- (11.1, 4.5); % K
	\draw[color=gray, dashed] (7.5, 5.4) -- (11.1, 5.4); % L
	\draw[color=gray, dashed] (7.5, 6) -- (11.1, 6); % M and N
	\draw[color=gray, dashed] (7.5, 7.2) -- (11.1, 7.2); % O
\end{tikzpicture}
\begin{tikzpicture}
	% Samples at x = 0, 4, and 6
	\node [circle, fill, scale=0.2,label=above right:{\small A}] at (0,0) {A};
	\draw (-0.4, -0.2) -- (0.4, -0.2) -- (0.4, -0.3) -- (-0.4, -0.3) -- (-0.4, -0.2);
	\draw (-0.1, -0.1) -- (-0.1, -0.4);
	\node [label=below:{\small $v$}] at (-0.1, -0.2) {};
	\node [circle, fill, scale=0.2,label=above right:{\small B}] at (4,0) {B};
	\draw (3.6, -0.2) -- (4.4, -0.2) -- (4.4, -0.3) -- (3.6, -0.3) -- (3.6, -0.2);
	\node [circle, fill, scale=0.2,label=above left:{\small C}] at (6,0) {C};
	\draw (5.6, -0.2) -- (6.4, -0.2) -- (6.4, -0.3) -- (5.6, -0.3) -- (5.6, -0.2);

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (11, 1.2); % D
	\draw[color=gray, dashed] (7.5, 1.5) -- (11, 1.5); % E and F
	\draw[color=gray, dashed] (7.5, 2.4) -- (11, 2.4); % G and H
	\draw[color=gray, dashed] (7.5, 3.1) -- (11, 3.1); % I and J
	\draw[color=gray, dashed] (7.5, 3.8) -- (11, 3.8);	% P and Q
	\draw[color=gray, dashed] (7.5, 4.5) -- (11, 4.5); % K
	\draw[color=gray, dashed] (7.5, 5.4) -- (11, 5.4); % L
	\draw[color=gray, dashed] (7.5, 6) -- (11, 6); % M and N
	\draw[color=gray, dashed] (7.5, 6.6) -- (11, 6.6);	% R
	\draw[color=gray, dashed] (7.5, 7.2) -- (11, 7.2); % O
\end{tikzpicture}
\caption{(Top) A realisation of the graph traversed by Hudson's algorithm started from a sample of three continuous chromosomes of length $L$ at time $t = 0$, and propagated until time $T$.
The MRCA on the genetic interval $[x, z)$ is reached at node L, while that on $[0, x)$ is reached at node O, which is a root as it carries only fully coalesced ancestral segments.
The non-ancestral segment $[x, y)$ above node K contributes to the rate of effective recombinations because it is trapped between ancestral segments.
The two columns titled ``Extant material" are the number of extant lineages and the total length of genetic material on which an effective recombination can take place during each time interval.
(Bottom) A corresponding realisation of a big ARG, which augments Hudson's algorithm by tracking nonancestral lineages.
The result is a simpler state space and dynamics, at the cost of extra nodes and edges, highlighted in red, which do not affect the local tree at any site.}
\label{hudson_vs_bigARG}
\end{figure}

\begin{figure}
\vspace{5em}
\caption{\label{fig-pedigree-and-arg}
The (unannotated) ARG is closely related to the pedigree
graph, where each ``individual'' node is replace by two ``genome''
nodes, and edges between parents and children are only included if there is
transfer of ancestral material. We usually only include nodes in an ARG
that affect our sample, so that, for example, individuals that are ancestral
to only one sampled genome are not included.}
\end{figure}

>>>>

\begin{figure}[ht]
\centering
% FIXME this is a quick nasty hack to make the figure a bit smaller and
% prevent flushing all the other figures to the end of the document.
\scalebox{0.5}{
\begin{tabular}{cc}
\begin{tikzpicture}
	% Samples at x = 0, 4, and 6
	\node [circle, fill, scale=0.2,label=above right:{\small A}] at (0,0) {A};
	\draw (-0.4, -0.2) -- (0.4, -0.2) -- (0.4, -0.3) -- (-0.4, -0.3) -- (-0.4, -0.2);
	\draw (-0.1, -0.1) -- (-0.1, -0.4);
	\node [label=below:{\small $v$}] at (-0.1, -0.2) {};
	\node [circle, fill, scale=0.2,label=above right:{\small B}] at (4,0) {B};
	\draw (3.6, -0.2) -- (4.4, -0.2) -- (4.4, -0.3) -- (3.6, -0.3) -- (3.6, -0.2);
	\node [circle, fill, scale=0.2,label=above left:{\small C}] at (6,0) {C};
	\draw (5.6, -0.2) -- (6.4, -0.2) -- (6.4, -0.3) -- (5.6, -0.3) -- (5.6, -0.2);

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (11.1, 1.2); % D
	\draw[color=gray, dashed] (7.5, 1.5) -- (11.1, 1.5); % E and F
	\draw[color=gray, dashed] (7.5, 2.4) -- (11.1, 2.4); % G and H
	\draw[color=gray, dashed] (7.5, 3.1) -- (11.1, 3.1); % I and J
	\draw[color=gray, dashed] (7.5, 4.5) -- (11.1, 4.5); % K
	\draw[color=gray, dashed] (7.5, 5.4) -- (11.1, 5.4); % L
	\draw[color=gray, dashed] (7.5, 6) -- (11.1, 6); % M and N
	\draw[color=gray, dashed] (7.5, 7.2) -- (11.1, 7.2); % O
\end{tikzpicture}
&
\begin{tikzpicture}
	% Samples at x = 0, 4, and 6
	\node [circle, fill, scale=0.2,label=above right:{\small A}] at (0,0) {A};
	\draw (-0.4, -0.2) -- (0.4, -0.2) -- (0.4, -0.3) -- (-0.4, -0.3) -- (-0.4, -0.2);
	\draw (-0.1, -0.1) -- (-0.1, -0.4);
	\node [label=below:{\small $v$}] at (-0.1, -0.2) {};
	\node [circle, fill, scale=0.2,label=above right:{\small B}] at (4,0) {B};
	\draw (3.6, -0.2) -- (4.4, -0.2) -- (4.4, -0.3) -- (3.6, -0.3) -- (3.6, -0.2);
	\node [circle, fill, scale=0.2,label=above left:{\small C}] at (6,0) {C};
	\draw (5.6, -0.2) -- (6.4, -0.2) -- (6.4, -0.3) -- (5.6, -0.3) -- (5.6, -0.2);

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (11, 1.2); % D
	\draw[color=gray, dashed] (7.5, 1.5) -- (11, 1.5); % E and F
	\draw[color=gray, dashed] (7.5, 2.4) -- (11, 2.4); % G and H
	\draw[color=gray, dashed] (7.5, 3.1) -- (11, 3.1); % I and J
	\draw[color=gray, dashed] (7.5, 3.8) -- (11, 3.8);	% P and Q
	\draw[color=gray, dashed] (7.5, 4.5) -- (11, 4.5); % K
	\draw[color=gray, dashed] (7.5, 5.4) -- (11, 5.4); % L
	\draw[color=gray, dashed] (7.5, 6) -- (11, 6); % M and N
	\draw[color=gray, dashed] (7.5, 6.6) -- (11, 6.6);	% R
	\draw[color=gray, dashed] (7.5, 7.2) -- (11, 7.2); % O
\end{tikzpicture}
\end{tabular}
}
\caption{(Top) A realisation of the graph traversed by Hudson's algorithm started from a sample of three continuous chromosomes of length $L$ at time $t = 0$, and propagated until time $T$.
The MRCA on the genetic interval $[x, z)$ is reached at node L, while that on $[0, x)$ is reached at node O, which is a root as it carries only fully coalesced ancestral segments.
The non-ancestral segment $[x, y)$ above node K contributes to the rate of effective recombinations because it is trapped between ancestral segments.
The two columns titled ``Extant material" are the number of extant lineages and the total length of genetic material on which an effective recombination can take place during each time interval.
(Bottom) A corresponding realisation of a big ARG, which augments Hudson's algorithm by tracking nonancestral lineages.
The result is a simpler state space and dynamics, at the cost of extra nodes and edges, highlighted in red, which do not affect the local tree at any site.}
\label{hudson_vs_bigARG}
\end{figure}

@@

We describe here an alternative formulation in which the
concepts are based on the diploid pedigree, which
is free from the limitations discussed in the previous section.
\cite{mathieson2020ancestry} and others
\citep{wakeley2012genetics,gusfield2014recombinatorics,speed2015naturereviewsgenetics}
have also suggested this approach, although without explicitly defining
the roles of nodes and the details of how ancestry
flows along edges.
\begin{figure}
\vspace{5em}
\caption{\label{fig-pedigree-and-arg}
The (unannotated) ARG is closely related to the pedigree
graph, where each ``individual'' node is replace by two ``genome''
nodes, and edges between parents and children are only included if there is
transfer of ancestral material. We usually only include nodes in an ARG
that affect our sample, so that, for example, individuals that are ancestral
to only one sampled genome are not included.}
\end{figure}
